### 2023-01-10 manuscript.rmd
small edits

Comparative approaches are frequently used in communication research, especially in journalism studies. Journalism studies are using the comparative approach [@hanusch:2019:C] to model journalistic culture [e.g. @esser:2013:C], news value [e.g. @burggraaff:2017:T;@wilke:2012], news flow [e.g. @wu:2000:SDI;@grasland:2019:I], among others. Around 40% of recent  comparative studies in journalism research are comparative content analysis [@hanusch:2019:C], which include news articles from various outlets and usually also from various countries. Then the included news articles are coded either manually or automatically. Another 10% of recent comparative studies in journalism research are surveys such as the Worlds of Journalism Study (WJS) [@hanusch:2017:CJC]. Although this article uses examples from journalism studies, the comparative approach is not bound to journalism studies and has been widely deployed in various subfields of communication. Comparative communication research is not exclusive to comparison across countries. Based on @de2017comparative’s classification, one can do comparisons across times, media, and other units [e.g. language regions within a multilingual country, see @vogler:2021:TNM]. In social media studies, for instance, it is common to compare content across sources [e.g. @zhao2019appealing]. The methodological issues discussed in this paper apply also to those cases.

@@

This study of ecological effects on individual behaviors has a long tradition in communication research. There are hypothesis-generating and hypothesis- testing comparative studies. An example of hypothesis-generating comparative research is <!-- redacted for ijoc  -->AUTHOR1 <!-- @chan2020combining -->, in which the sentiment profiles of terrorism coverage from Muslim- and Christian-majority countries were visualized. Hypothesis-generating comparative study, however, is rare. As we see from the example hypotheses above, most of these studies propose hypotheses to test for ecological effects. In this context usually traditional (frequentist) hypothesis testing approaches were used: @esser:2013:C lumped multiple outlets from the same country together and then used univariate analysis of variance to test for the differences in the proportion of opinion-orienting articles across countries. In @wu:2000:SDI, multiple stepwise regression was used. These approaches violate the underlying independence of observations assumption of linear regression (See non-atomicity below). It highlights the fact that comparative research introduces a feature that researchers usually overlook: media contents are clustered in a multilevel structure. A news article is nested within its media outlet and its media outlet is in turn nested within its country. Such data structure brings two problems: non-atomicity and non-stochasticity. And specifically, we propose two solutions: multilevel modeling and Bayesian approach.

The goal of this paper is twofold. We use examples from journalism studies as the most critical cases in communication science to highlight the strength of multilevel models and the advantages of Bayesian models. In our paper, we first discuss these two issues and then suggest Bayesian multilevel regression as the most defensible option. Using the openly available WJS and *useNews* datasets, we then demonstrate how to apply a Bayesian approach for the analysis of comparative data.

@@

In comparative research, it is easy to assume that macro-level independent variables (e.g. Anglo-American systems) could be analyzed at the same level as the micro-level dependent variable (e.g. the degree of opinion-orientation). The manifestation of this assumption is to enter macro-level variables as independent variables in multiple regression analysis and regress them against a micro-level dependent variable. Suppose one wants to study how the democratic performance of a country affects journalists’ perceived professional autonomy using the WJS data [Wave 1: @reich:2013:DJP; Wave 2: @hamada:2021:DJA]. The relationship can be expressed in the following regression equation: let $y_{i}$ be the dependent variable at the micro-level (journalist-level: perceived professional autonomy) and $x_{i}$ be the independent variable at the macro-level (country-level: democratic performance). Suppose there are $m$ journalists  where $i = 1,2,\ldots,m$.

@@

This method is the so-called "disaggregation approach". From an epistemic standpoint, searching for ecological effect by studying the value of slope ($\beta_{1}$) leads to the so-called atomistic fallacy [@hox2017multilevel]. From a statistical standpoint, this approach violates the underlying independence assumption. Almost all frequentist statistical tests assume that the observations are independent of each other: $y_{i}$ given $x_{i}$ are independently identically distributed (i.i.d.). Journalists from the same country are hardly i.i.d.: they are subjected to being governed by the same government and exposed to the same journalistic culture. Also, the way communication researchers survey the journalists resembles (nonrandom) cluster sampling rather than pure random sampling: one usually starts from a specific country and then surveys journalists from it. The choice of countries is almost always not randomly selected. This sampling approach makes the independence assumption even more fragile. Previous simulation studies have shown that ignoring this dependence can lead to false-positive associations [e.g. @clarke:2008:W;@chen:2012:IIL].

@@

[^indieerror]: In general, the non-atomicity alone only influences the inference (e.g. p-values, confidence intervals) by increasing the chance of Type I errors. The magnitude of the regression coefficient should still be consistent. Another approach, which we are not going to discuss in detail, is to use the clustered-robust standard error. For a survey of the technique, see @wooldridge:2003:CSM. The technique can only address the non-atomicity issue alone, but not non-stochasticity.

@@

Using this aggregation method, the unit of analysis effectively switches from journalist to country. This method is useful when $x$ is the only independent variable. It is not technically committing the atomistic fallacy, when one uses the value of slope as the evidence for an ecological effect. However, this method still has important drawbacks. First, it cannot be used for data with more than two levels. In those cases, the lower-level predictors were lumped together and in effect, assumed to be homogeneous and discarded. Second, this method discards a massive amount of information. It is useful only for analysis of a dependent micro-level variable with a reasonable aggregation function (e.g. counting a binary variable). For numerical variables, aggregation functions such as taking a mean or median cannot capture the spread of the micro-level variable [@bryk:1988:H]. Also, the effective sample size is reduced from the number of journalists ($m$) to the number of countries ($n$). Third, if one specifies hypotheses at the journalist-level but bases one’s conclusion on the aggregated analysis, this is a risk of interpreting the macro-level association wrongly at the micro-level, otherwise known as ecological fallacy. Nonetheless, this aggregation method, although inflexible, is still useful when the number of groups (e.g. $n$) is large. It is also useful to collapse micro-levels (e.g. article-level) that are not useful in answering one’s research questions.

@@

The advantage of using multilevel modeling lies in its flexibility in handling multi-level data. Suppose we also want to consider the clustering of journalists around $o$ different media organizations in the above example and $j$-th media media organization of the $i$-th journalist, where $j = 1,\ldots, o$ and if such data were available. [^WJSOUTLET] The multilevel regression equations are rewritten as:

@@

This model is called the varying slope model and is useful to establish robust estimation of effect (in this case $\gamma_{10}$) across all countries. In this paper, we will not go into detail of this kind of model. For an example of its application in comparative communication research, see @barnidge:2018:SMS.

@@

@stegmueller:2013:HMC demonstrates that MLE for multilevel modeling is associated with shrinkage (reduction of standard error, i.e. more false positives) and the shrinkage is more severe when the number of macro-level units (e.g. countries) is small. @stegmueller:2013:HMC proposes to use Bayesian analysis as a robust alternative [see counterarguments from @elff:2020:MAF]. Although less restrictive methods such as restricted maximum likelihood (REML) have been demonstrated to remediate the shrinkage issue of MLE [@elff:2020:MAF], we still agree with Stegmueller’s proposal for theoretical reasons [@western:1994:BIC].

@@

Most of the comparative content analytic studies, for example, often collect all available content data from a bunch of selected media outlets. In contrast to cluster sampling where media outlets are sampled randomly from a sampling frame of all media outlets, these studies are collecting the entire population of observations from some nonprobailistically selected media outlets. In these census-like situations, there is no way to get more data unless the scope of these studies is changed [@berk:1995:SIA]. It is especially true for modern large-N studies using automated content analytic techniques. @burggraaff:2017:T, for instance, "collected **all** available news items from **a selection of major Dutch news outlets**, both online and print" (p. 6, emphasis added) and that amounted to 762,095 articles from 9 outlets in the period of 2014--2015. In that study, one could only get more data by including more media outlets or widening the time window [^conventionalcontentanalysis]. Unlike a (theoretically) repeatable survey like WJS, one can randomly select more journalists from the sampling frame of all journalists in the respective country.

[^conventionalcontentanalysis]: Comparative content analytic studies can also be performed without this census-like approach. The traditional approach of making probabilistic sample from all units can also be used [@krippendorff2018content]. As there is an explicit sampling frame of all units, the population of content about which the sample is making inference from is known in this case. Any further inference made beyond the population is speculative.

@@

Without repeated experiment, these probabilities quantify our certainty on how plausible the paper is being accepted, given the current available data. We revise our old beliefs (or **prior**, $p(\theta)$) with the new data ($X$) and form our revised belief (or **posterior**, $p(\theta|X)$). This can be summarized in the following equation [@gelman2020bayesian]:

@@

[^mcmcexplanation]: How concretely MCMC can faciliate the estimation of likelihood function, given the data and prior, is too technical to be included in this paper. Readers can get an intuition of how it works in Chapter 8 of @mcelreath2020statistical. For how the method works in regression models, Chapters 9, 10, and 12 of @mcelreath2020statistical. For a comparative perspective between Bayesian and frequentist approaches, please refer to @samaniego2010comparison.

R interfaces to Stan (the probabilistic programming language for conducting MCMC), such as *brms* [@burkner2017advanced], can enable the derivation of the likelihood function (Part 2). But still, it is important to be mindful that Bayesian analysis is much more computational intensive than methods such as MLE. Our benchmark suggests that Bayesian analysis needs at least 100 times more running time than MLE.

Part 3 (Prior) is arguably the most controversial part of Bayesian analysis. In the *IJOC* example, we can select a reasonable prior (or **informative** prior) of 24% from published information. Finding previous studies for an informative prior is the logical first thing to do. @keating:2019:W show that one in every seven communication research papers published in major communication journals was a form of replication attempt. For these one seventh of communication research, there should be previous studies available to base one’s informative prior non-controversially. For the other six sevenths, one might not have any information to set an informative prior. One option is to consult experts or to make an educated guess. Expert elicitation is a way to probe how the experts in the field think about the hypotheses. A standardized protocol for expert elicitation is available [@hanea:2017:IDE] and there are many software tools available to facilitate the process [^experttools]. But one person's expert opinion could be another person's wishful thinking. And this perceived subjectivity of specifying priors by experts’ judgment attracts wide-spread criticism from both statisticians [e.g. @efron1986isn] and social scientists [e.g. @elff:2020:MAF].

@@

By comparing the point and interval estimates using different approaches, point estimates are remarkably similar. Bayesian multilevel regression gave a wider 95% HDI than confidence intervals from frequentist methods. It is like previous studies [@stegmueller:2013:HMC; @elff:2020:MAF]. While the frequentist confidence intervals and the Bayesian credible interval seem to be similar, they are from a philosophical point of view totally different and lead to different forms of inference. The Bayesian approach allows us to say that *given the observed data, the true value of the standardized regression coefficient has 95% probability of falling within  0.13 and 0.34*. In contrast, the frequentist confidence interval only allows us to say that *when computing a confidence interval from the same type of data in repeated studies (if possible), 95% of the confidence intervals will include the true value of the standardized regression coefficient*. One of the misconceptions of frequentist confidence intervals is that they can be interpreted in a Bayesian way as described above [@morey:2015]. This example illustrates that a Bayesian approach allows us to interpret the results in a more intuitive way.

@@

We used the useNews dataset [@puschmann:2020] to demonstrate how to perform a Bayesian multilevel regression analysis on a nonstochastic comparative dataset. The analysis was guided by a classic question in news value research: Does the distance between China and the host country of a media outlet increase the frequency of China coverage? In other words, we hypothesized that there is an ecological effect between a closer distance to China and an increased frequency of China coverage. Several distance measures were used to formulate our pre-registered hypotheses<!-- : physical distance between two capitals, trade volume between two countries, and cultural distance between two societies -->. Due to limited space, only the analysis of trade volume between two countries is displayed here. For other analyses, please refer to the online appendix.

For demonstrative purposes, our analysis was separated into two levels: outlet-level analysis (here) and article-level analysis (in the Online Appendix). The outlet-level analysis is an aggregated version of article-level analysis. It is used to demonstrate the flexibility of Bayesian analysis to handle two-level data with a very small data size. The article-level analysis is used to demonstrate the analysis of a massive three-level dataset.

>>>>

Comparative approaches are frequently used in communication research, especially in journalism studies. Journalism studies are using the comparative approach [@hanusch:2019:C] to model journalistic culture [e.g. @esser:2013:C], news value [e.g. @burggraaff:2017:T;@wilke:2012], news flow [e.g. @wu:2000:SDI;@grasland:2019:I], among others. Around 40% of recent comparative studies in journalism research are comparative content analysis [@hanusch:2019:C], which include news articles from various outlets and usually also from various countries. Then the included news articles are coded either manually or automatically. Another 10% of recent comparative studies in journalism research are surveys such as the Worlds of Journalism Study (WJS) [@hanusch:2017:CJC]. Although this article uses examples from journalism studies, the comparative approach is not bound to journalism studies and has been widely deployed in various subfields of communication. Comparative communication research is not exclusive to comparison across countries. Based on @de2017comparative’s classification, one can do comparisons across times, media, and other units [e.g. language regions within a multilingual country, see @vogler:2021:TNM]. In social media studies, for instance, it is common to compare content across sources [e.g. @zhao2019appealing]. The methodological issues discussed in this paper also apply to those cases.

@@

This study of ecological effects on individual behaviors has a long tradition in communication research. There are hypothesis-generating and hypothesis-testing comparative studies. An example of hypothesis-generating comparative research is <!-- redacted for ijoc  -->AUTHOR1 <!-- @chan2020combining -->, in which the sentiment profiles of terrorism coverage from Muslim- and Christian-majority countries were visualized. Hypothesis-generating comparative study, however, is rare. As we see from the example hypotheses above, most of these studies propose hypotheses to test for ecological effects. In this context usually traditional (frequentist) hypothesis testing approaches were used: @esser:2013:C lumped multiple outlets from the same country together and then used univariate analysis of variance to test for the differences in the proportion of opinion-orienting articles across countries. In @wu:2000:SDI, multiple stepwise regression was used. These approaches violate the underlying independence of observations assumption of linear regression (See non-atomicity below). It highlights the fact that comparative research introduces a feature that researchers usually overlook: media contents are clustered in a multilevel structure. A news article is nested within its media outlet and its media outlet is, in turn, nested within its country. Such data structure brings two problems: non-atomicity and non-stochasticity. And specifically, we propose two solutions: multilevel modeling and Bayesian approach.

The goal of this paper is twofold. We use examples from journalism studies as the most critical cases in communication science to highlight the strength of multilevel models and the advantages of Bayesian models. In our paper, we first discuss these two issues and then suggest Bayesian multilevel regression as the most defensible option. Using the openly available WJS and *useNews* datasets, we then demonstrate how to apply a Bayesian approach to analyze comparative data.

@@

In comparative research, it is easy to assume that macro-level independent variables (e.g., Anglo-American systems) could be analyzed at the same level as the micro-level dependent variable (e.g., the degree of opinion-orientation). The manifestation of this assumption is to enter macro-level variables as independent variables in multiple regression analysis and regress them against a micro-level dependent variable. Suppose one wants to study how the democratic performance of a country affects journalists’ perceived professional autonomy using the WJS data [Wave 1: @reich:2013:DJP; Wave 2: @hamada:2021:DJA]. The relationship can be expressed in the following regression equation: let $y_{i}$ be the dependent variable at the micro-level (journalist-level: perceived professional autonomy) and $x_{i}$ be the independent variable at the macro-level (country-level: democratic performance). Suppose there are $m$ journalists  where $i = 1,2,\ldots,m$.

@@

This method is the so-called "disaggregation approach". From an epistemic standpoint, searching for ecological effect by studying the value of slope ($\beta_{1}$) leads to the so-called atomistic fallacy [@hox2017multilevel]. From a statistical standpoint, this approach violates the underlying independence assumption. Almost all frequentist statistical tests assume that the observations are independent of each other: $y_{i}$ given $x_{i}$ are independently identically distributed (i.i.d.). Journalists from the same country are hardly i.i.d.: they are subjected to being governed by the same government and exposed to the same journalistic culture. Also, the way communication researchers survey the journalists resembles (nonrandom) cluster sampling rather than pure random sampling: one usually starts from a specific country and then surveys journalists from it. The choice of countries is almost always not randomly selected. This sampling approach makes the independence assumption even more fragile. Previous simulation studies have shown that ignoring this dependence can lead to false-positive associations [e.g., @clarke:2008:W;@chen:2012:IIL].

@@

[^indieerror]: In general, the non-atomicity alone only influences the inference (e.g., p-values, confidence intervals) by increasing the chance of Type I errors. The magnitude of the regression coefficient should still be consistent. Another approach, which we are not going to discuss in detail, is to use the clustered-robust standard error. For a survey of the technique, see @wooldridge:2003:CSM. The technique can only address the non-atomicity issue alone, but not non-stochasticity.

@@

Using this aggregation method, the unit of analysis effectively switches from journalist to country. This method is useful when $x$ is the only independent variable. It is not technically committing the atomistic fallacy, when one uses the value of slope as the evidence for an ecological effect. However, this method still has important drawbacks. First, it cannot be used for data with more than two levels. In those cases, the lower-level predictors were lumped together and in effect, assumed to be homogeneous and discarded. Second, this method discards a massive amount of information. It is useful only for the analysis of a dependent micro-level variable with a reasonable aggregation function (e.g. counting a binary variable). For numerical variables, aggregation functions such as taking a mean or median cannot capture the spread of the micro-level variable [@bryk:1988:H]. Also, the effective sample size is reduced from the number of journalists ($m$) to the number of countries ($n$). Third, if one specifies hypotheses at the journalist-level but bases one’s conclusion on the aggregated analysis, this is a risk of interpreting the macro-level association wrongly at the micro-level, otherwise known as ecological fallacy. Nonetheless, this aggregation method, although inflexible, is still useful when the number of groups (e.g. $n$) is large. It is also useful to collapse micro-levels (e.g. article-level) that are not useful in answering one’s research questions.

@@

The advantage of using multilevel modeling lies in its flexibility in handling multilevel data. Suppose we also want to consider the clustering of journalists around $o$ different media organizations in the above example and $j$-th media organization of the $i$-th journalist, where $j = 1,\ldots, o$ and if such data were available. [^WJSOUTLET] The multilevel regression equations are rewritten as:

@@

This model is called the varying slope model and is useful to establish robust estimation of effect (in this case $\gamma_{10}$) across all countries. In this paper, we will not go into detail about this kind of model. For an example of its application in comparative communication research, see @barnidge:2018:SMS.

@@

@stegmueller:2013:HMC demonstrates that MLE for multilevel modeling is associated with shrinkage (reduction of standard error, i.e. more false positives), and the shrinkage is more severe when the number of macro-level units (e.g., countries) is small. @stegmueller:2013:HMC proposes to use Bayesian analysis as a robust alternative [see counterarguments from @elff:2020:MAF]. Although less restrictive methods such as restricted maximum likelihood (REML) have been demonstrated to remediate the shrinkage issue of MLE [@elff:2020:MAF], we still agree with Stegmueller’s proposal for theoretical reasons [@western:1994:BIC].

@@

Most of the comparative content analytic studies, for example, often collect all available content data from a bunch of selected media outlets. In contrast to cluster sampling where media outlets are sampled randomly from a sampling frame of all media outlets, these studies collect the entire population of observations from some non-probabilistically selected media outlets. In these census-like situations, there is no way to get more data unless the scope of these studies is changed [@berk:1995:SIA]. It is especially true for modern large-N studies using automated content analytic techniques. @burggraaff:2017:T, for instance, "collected **all** available news items from **a selection of major Dutch news outlets**, both online and print" (p. 6, emphasis added) and that amounted to 762,095 articles from 9 outlets in the period of 2014--2015. In that study, one could only get more data by including more media outlets or widening the time window [^conventionalcontentanalysis]. Unlike a (theoretically) repeatable survey like WJS, one can randomly select more journalists from the sampling frame of all journalists in the respective country.

[^conventionalcontentanalysis]: Comparative content analytic studies can also be performed without this census-like approach. The traditional approach of making a probabilistic sample from all units can also be used [@krippendorff2018content]. As there is an explicit sampling frame of all units, the population of content about which the sample is making inference from is known in this case. Any further inference made beyond the population is speculative.

@@

Without repeated experiments, these probabilities quantify our certainty on how plausible the paper is being accepted, given the currently available data. We revise our old beliefs (or **prior**, $p(\theta)$) with the new data ($X$) and form our revised belief (or **posterior**, $p(\theta|X)$). This can be summarized in the following equation [@gelman2020bayesian]:

@@

[^mcmcexplanation]: How concretely MCMC can facilitate the estimation of likelihood function, given the data and prior, is too technical to be included in this paper. Readers can get an intuition of how it works in Chapter 8 of @mcelreath2020statistical. For how the method works in regression models, Chapters 9, 10, and 12 of @mcelreath2020statistical. For a comparative perspective between Bayesian and frequentist approaches, please refer to @samaniego2010comparison.

R interfaces to Stan (the probabilistic programming language for conducting MCMC), such as *brms* [@burkner2017advanced], can enable the derivation of the likelihood function (Part 2). But still, it is important to be mindful that Bayesian analysis is much more computationally intensive than methods such as MLE. Our benchmark suggests that Bayesian analysis needs at least 100 times more running time than MLE.

Part 3 (Prior) is arguably the most controversial part of Bayesian analysis. In the *IJOC* example, we can select a reasonable prior (or **informative** prior) of 24% from published information. Finding previous studies for an informative prior is the logical first thing to do. @keating:2019:W show that one in every seven communication research papers published in major communication journals was a form of replication attempt. For these one-seventh of communication research, there should be previous studies available to base one’s informative prior non-controversially. For the other six-sevenths, one might not have any information to set an informative prior. One option is to consult experts or to make an educated guess. Expert elicitation is a way to probe how the experts in the field think about the hypotheses. A standardized protocol for expert elicitation is available [@hanea:2017:IDE] and there are many software tools available to facilitate the process [^experttools]. But one person's expert opinion could be another person's wishful thinking. And this perceived subjectivity of specifying priors by experts’ judgment attracts widespread criticism from both statisticians [e.g. @efron1986isn] and social scientists [e.g. @elff:2020:MAF].

@@

By comparing the point and interval estimates using different approaches, point estimates are remarkably similar. Bayesian multilevel regression gave a wider 95% HDI than confidence intervals from frequentist methods. It is like previous studies [@stegmueller:2013:HMC; @elff:2020:MAF]. While the frequentist confidence intervals and the Bayesian credible interval seem to be similar, they are totally different from a philosophical point of view and lead to different forms of inference. The Bayesian approach allows us to say that *given the observed data, the true value of the standardized regression coefficient has 95% probability of falling within  0.13 and 0.34*. In contrast, the frequentist confidence interval only allows us to say that *when computing a confidence interval from the same type of data in repeated studies (if possible), 95% of the confidence intervals will include the true value of the standardized regression coefficient*. One of the misconceptions of frequentist confidence intervals is that they can be interpreted in a Bayesian way as described above [@morey:2015]. This example illustrates that a Bayesian approach allows us to interpret the results in a more intuitive way.

@@

We used the useNews dataset [@puschmann:2020] to demonstrate how to perform a Bayesian multilevel regression analysis on a non-stochastic comparative dataset. The analysis was guided by a classic question in news value research: Does the distance between China and the host country of a media outlet increase the frequency of China coverage? In other words, we hypothesized that there is an ecological effect between a closer distance to China and an increased frequency of China coverage. Several distance measures were used to formulate our pre-registered hypotheses<!-- : physical distance between two capitals, trade volume between two countries, and cultural distance between two societies -->. Due to limited space, only the analysis of trade volume between two countries is displayed here. For other analyses, please refer to the online appendix.

For demonstrative purposes, our analysis was separated into two levels: outlet-level analysis (here) and article-level analysis (in the Online Appendix). The outlet-level analysis is an aggregated version of article-level analysis. It is used to demonstrate the flexibility of Bayesian analysis to handle two-level data with very small data size. The article-level analysis is used to demonstrate the analysis of a massive three-level dataset.
